---
title: "Final Project 714"
author: "Bridget_Chukwu"
date: "July 30, 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploring the pattern of macroeconomic variables on different North American countries through interactive Data Visualization

#### Data cleaning and preprocessing to handle missing values, outliers, and inconsistencies
```{r}

library(readxl)
getwd()
data<-read_excel("C:\\Users\\bridget.chukwu\\Documents\\Summer Course STAT\\STAT 714\\Final Project_714\\MacroVariables Data.xls")
head(data)
dim(data) # List the names of columns in the dataset
```

### Cleaning the data 
```{r}
# Remove unneeded columns and columns with zero values from data dataset
data <- data[, !(names(data) %in% c("Time" , "Entity", "lnGDP", "lnGovEx", "lnInf", "lnInt", "lnTax", "lnUnem", "CAN", "CRI", "DOM", "HND", "MEX", "PAN"))]
head(data)
dim(data)
colnames(data)
###missing value
sum(is.na(data)) # input the formula to detect any missing value
#no missing value found

### 

```

 
### Outliers detection

```{r}
# Standardize the numeric columns to z-scores
z_scores <- apply(data[, sapply(data, is.numeric)], 2, function(x) (x - mean(x)) / sd(x))

# Identify rows with any z-score greater than 3 (considered outliers)
outlier_rows <- apply(z_scores, 1, function(row) any(abs(row) > 3))

# Create a dataset without the outliers
cleaned_data <- data[!outlier_rows, ]

# Ensure at least one record per country remains
country_column <- "countrycode"  
unique_countries <- unique(data[[country_column]])

# Loop over countries to check if they are represented
for (country in unique_countries) {
  if (!country %in% cleaned_data[[country_column]]) {
    # If a country is not present in the cleaned data, retain one row from the original data
    row_to_keep <- My_data[My_data[[country_column]] == country, ][1, ]
    cleaned_data <- rbind(cleaned_data, row_to_keep)
  }
}

# Check the remaining dimension after removing outliers
dim(cleaned_data)

# Optionally, reassign cleaned data to original variable
final_data <- cleaned_data



```

### Data visualization - Scatter Plot

## Box Plot
```{r}
library(plotly)
library(ggplot2)
# Data visualization - 
boxplot(final_data$`Unem`, ylab = "Unem") #create boxplot for Unmeployment
out_Un<- boxplot.stats(final_data$`Unem`)$out # show the statistics from the boxplot and detect the outliers in the dataset
outUn_ind<- which(final_data$`Unem` %in% c(out_Un)) #create a variable outUn_ind with the detected outliers
outUn_ind
boxplot(final_data$`Unem`, ylab = "Unem", main = "Unem")#create a boxplot
mtext(paste("outliers: ", paste(out_Un, collapse = ", "))) #show the outliers value on the boxplot
```

```{r}
library(plotly)
library(ggplot2)

# Data visualization - Boxplot for Tax
boxplot(final_data$`Tax`, ylab = "Tax") # Create boxplot for Tax
out_Tax <- boxplot.stats(final_data$`Tax`)$out # Detect the outliers in the dataset
outTax_ind <- which(final_data$`Tax` %in% c(out_Tax)) # Create a variable outTax_ind with the detected outliers
outTax_ind

boxplot(final_data$`Tax`, ylab = "Tax", main = "Tax") # Create a boxplot for Tax
mtext(paste("outliers: ", paste(out_Tax, collapse = ", "))) # Show the outliers value on the boxplot

```


```{r}
library(plotly)
library(ggplot2)

# Data visualization - Boxplot for Inf
boxplot(final_data$`Inf`, ylab = "Inf") # Create boxplot for Inf
out_Inf <- boxplot.stats(final_data$`Inf`)$out # Detect the outliers in the dataset
outInf_ind <- which(final_data$`Inf` %in% c(out_Inf)) # Create a variable outInf_ind with the detected outliers
outInf_ind

boxplot(final_data$`Inf`, ylab = "Inf", main = "Inf") # Create a boxplot for Inf
mtext(paste("outliers: ", paste(out_Inf, collapse = ", "))) # Show the outliers value on the boxplot

```

```{r}
library(plotly)
library(ggplot2)

# Data visualization - Boxplot for GDP
boxplot(final_data$`GDP`, ylab = "GDP") # Create boxplot for GDP
out_GDP <- boxplot.stats(final_data$`GDP`)$out # Detect the outliers in the dataset
outGDP_ind <- which(final_data$`GDP` %in% c(out_GDP)) # Create a variable outGDP_ind with the detected outliers
outGDP_ind

boxplot(final_data$`GDP`, ylab = "GDP", main = "GDP") # Create a boxplot for GDP
mtext(paste("outliers: ", paste(out_GDP, collapse = ", "))) # Show the outliers value on the boxplot

```

```{r}
# Data visualization - GDP by Country
gdp_comparison <- final_data %>%
  plot_ly(x = ~countrycode, y = ~GDP, color = ~countrycode, type = "box") %>%
  layout(title = "GDP per capita Comparison by Country",
         xaxis = list(title = "Country"),
         yaxis = list(title = "GDP"),
         showlegend = FALSE)

gdp_comparison

```


```{r}
# Function to identify outlier indices based on IQR
outlier_indices <- function(x) {
  Q1 <- quantile(x, 0.25, na.rm = TRUE)
  Q3 <- quantile(x, 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  which(x < (Q1 - 1.5 * IQR) | x > (Q3 + 1.5 * IQR))
}

# Find outlier indices for each variable
outliers_int <- outlier_indices(final_data$Int)
outliers_gdp <- outlier_indices(final_data$GDP)
outliers_tax <- outlier_indices(final_data$Tax)
outliers_unem <- outlier_indices(final_data$Unem)

# Combine all outlier indices and remove duplicates
all_outliers <- unique(c(outliers_int, outliers_gdp, outliers_tax, outliers_unem))

# Remove rows with outliers from the dataset
final_data_clean <- final_data[-all_outliers, ]

# Plotting box plots using the cleaned data
boxplot(final_data_clean$Int, main = "Interest Rate (Cleaned)", ylab = "Interest Rate")
boxplot(final_data_clean$GDP, main = "GDP (Cleaned)", ylab = "GDP")
boxplot(final_data_clean$Tax, main = "Tax (Cleaned)", ylab = "Tax")
boxplot(final_data_clean$Unem, main = "Unemployment (Cleaned)", ylab = "Unemployment")


```

```{r}
# Data visualization - Int (Interest) by Country
int_comparison <- final_data %>%
  plot_ly(x = ~countrycode, y = ~Int, color = ~countrycode, type = "box") %>%
  layout(title = "Interest rate Comparison by Country",
         xaxis = list(title = "Country"),
         yaxis = list(title = "Interest Rate"),
         showlegend = FALSE)

int_comparison

# Data visualization - Unem (Unemployment) by Country
unem_comparison <- final_data %>%
  plot_ly(x = ~countrycode, y = ~Unem, color = ~countrycode, type = "box") %>%
  layout(title = "Unemployment rate Comparison by Country",
         xaxis = list(title = "Country"),
         yaxis = list(title = "Unemployment Rate"),
         showlegend = FALSE)

unem_comparison

# Data visualization - Tax rate by Country
tax_comparison <- final_data %>%
  plot_ly(x = ~countrycode, y = ~Tax, color = ~countrycode, type = "box") %>%
  layout(title = "Tax Comparison by Country",
         xaxis = list(title = "Country"),
         yaxis = list(title = "Tax"),
         showlegend = FALSE)

tax_comparison

```



## Bar Chart
#### Visualisation of the Aggregate of different variables by countries from 2004 - 2022
```{r}
library(plotly)
library(dplyr)

# Calculate the sum of Interest rate for each country
sum1 <- aggregate(Int ~ countrycode, data = final_data, FUN = sum)

# Sort the data by the sum of Int value in descending order
top_7_coun <- head(sum1[order(-sum1$Int), ], 7)

# Convert Int to factor to ensure the x-axis is discrete for bar chart
top_7_coun$countrycode <- factor(top_7_coun$countrycode)

fig1 <- plot_ly(data = top_7_coun, x = ~countrycode, y = ~Int,
                text = ~Int, textfont = list(color = 'rgb(101, 67, 33)'), textposition = 'inside', 
                type = "bar", marker = list(color = 'rgb(225, 253, 208)', 
                                            line = list(color = 'rgb(101, 67, 33)', 
                                                        width = 2))) %>%
   layout(title = "<b>Aggregated Interest rate values by Countries</b>",
         plot_bgcolor = 'white',
         xaxis = list(title = "<b>Countries</b>",
           zerolinecolor = 'ffff',
           zerolinewidth = 2,
           gridcolor = 'ffff'),
         yaxis = list(title = "<b>Agg.Int</b>",
           zerolinecolor = '#CD6600',
           zerolinewidth = 2,
           gridcolor = 'ffff'))

fig1

```


```{r}
library(plotly)
library(dplyr)

# Calculate the sum of Unemployment for each country
sum1 <- aggregate(Unem ~ countrycode, data = final_data, FUN = sum)

# Sort the data by the sum of  Unemployment  in descending order
top_7_coun <- head(sum1[order(-sum1$Unem), ], 7)

# Convert  Unemployment to factor to ensure the x-axis is discrete for bar chart
top_7_coun$countrycode <- factor(top_7_coun$countrycode)

fig1 <- plot_ly(data = top_7_coun, x = ~countrycode, y = ~Unem,
                text = ~Unem, textfont = list(color = 'rgb(101, 67, 33)'), textposition = 'inside', 
                type = "bar", marker = list(color = 'rgb(225, 253, 208)', 
                                            line = list(color = 'rgb(101, 67, 33)', 
                                                        width = 2))) %>%
   layout(title = "<b>Aggregated Unemployment values by Countries</b>",
         plot_bgcolor = 'white',
         xaxis = list(title = "<b>Countries</b>",
           zerolinecolor = 'ffff',
           zerolinewidth = 2,
           gridcolor = 'ffff'),
         yaxis = list(title = "<b>Agg.Unem</b>",
           zerolinecolor = '#CD6600',
           zerolinewidth = 2,
           gridcolor = 'ffff'))

fig1
```

```{r}

# Calculate the sum of Tax for each country
sum1 <- aggregate(Tax ~ countrycode, data = final_data, FUN = sum)

# Sort the data by the sum of Tax in descending order
top_7_coun <- head(sum1[order(-sum1$Tax), ], 7)

# Convert countrycode to factor to ensure the x-axis is discrete for the bar chart
top_7_coun$countrycode <- factor(top_7_coun$countrycode)

# Plot the data using plotly
fig1 <- plot_ly(data = top_7_coun, x = ~countrycode, y = ~Tax,
                text = ~Tax, textfont = list(color = 'rgb(101, 67, 33)'), textposition = 'inside', 
                type = "bar", marker = list(color = 'rgb(225, 253, 208)', 
                                            line = list(color = 'rgb(101, 67, 33)', 
                                                        width = 2))) %>%
   layout(title = "<b>Aggregated Tax  by Countries</b>",
         plot_bgcolor = 'white',
         xaxis = list(title = "<b>Countries</b>",
           zerolinecolor = 'ffff',
           zerolinewidth = 2,
           gridcolor = 'ffff'),
         yaxis = list(title = "<b>Agg.Tax</b>",
           zerolinecolor = '#CD6600',
           zerolinewidth = 2,
           gridcolor = 'ffff'))

fig1

```

```{r}

# Calculate the sum of GDP for each country
sum1 <- aggregate(GDP ~ countrycode, data = final_data, FUN = sum)

# Sort the data by the sum of Tax in descending order
top_7_coun <- head(sum1[order(-sum1$GDP), ], 7)

# Convert countrycode to factor to ensure the x-axis is discrete for the bar chart
top_7_coun$countrycode <- factor(top_7_coun$countrycode)

# Plot the data using plotly
fig1 <- plot_ly(data = top_7_coun, x = ~countrycode, y = ~GDP,
                text = ~GDP, textfont = list(color = 'rgb(101, 67, 33)'), textposition = 'inside', 
                type = "bar", marker = list(color = 'rgb(225, 253, 208)', 
                                            line = list(color = 'rgb(101, 67, 33)', 
                                                        width = 2))) %>%
   layout(title = "<b>Aggregated GDP  by Countries</b>",
         plot_bgcolor = 'white',
         xaxis = list(title = "<b>Countries</b>",
           zerolinecolor = 'ffff',
           zerolinewidth = 2,
           gridcolor = 'ffff'),
         yaxis = list(title = "<b>Agg.Tax</b>",
           zerolinecolor = '#CD6600',
           zerolinewidth = 2,
           gridcolor = 'ffff'))

fig1
```

## Scattered Plot
```{r}
# Load required library
library(plotly)

# Scatter plot for GDP vs Inf
fig_1 <- plot_ly(
  data = final_data, # Specify the data source
  x = ~final_data$`GDP`, # Set the X-axis data from the "GDP" column
  y = ~final_data$`Inf`, # Set the Y-axis data from the "Inf" (Inflation) column
  marker = list(
    size = 5, # Set the size of the markers
    color = "turquoise", # Set the color of the markers
    line = list(color = "gray0", width = 2) # Set the color and width of the marker borders
  )
)
fig_1 <- fig_1 %>% layout(
  title = "Change in GDP vs. Inflation", # Set the plot title
  yaxis = list(zeroline = FALSE, title = "Inflation (% change)"), # Customize the Y-axis properties
  xaxis = list(zeroline = FALSE, title = "GDP (% change)") # Customize the X-axis properties
)
fig_1 # Display the scatter plot

# Scatter plot for GDP vs Unem
fig_2 <- plot_ly(
  data = final_data, # Specify the data source
  x = ~final_data$`GDP`, # Set the X-axis data from the "GDP" column
  y = ~final_data$`Unem`, # Set the Y-axis data from the "Unem" (Unemployment) column
  marker = list(
    size = 5, # Set the size of the markers
    color = "blue", # Set the color of the markers
    line = list(color = "gray0", width = 2) # Set the color and width of the marker borders
  )
)
fig_2 <- fig_2 %>% layout(
  title = "Change in GDP vs. Unemployment", # Set the plot title
  yaxis = list(zeroline = FALSE, title = "Unemployment (% change)"), # Customize the Y-axis properties
  xaxis = list(zeroline = FALSE, title = "GDP (% change)") # Customize the X-axis properties
)
fig_2 # Display the scatter plot

# Scatter plot for GDP vs Int
fig_3 <- plot_ly(
  data = final_data, # Specify the data source
  x = ~final_data$`GDP`, # Set the X-axis data from the "GDP" column
  y = ~final_data$`Int`, # Set the Y-axis data from the "Int" (Interest Rate) column
  marker = list(
    size = 5, # Set the size of the markers
    color = "green", # Set the color of the markers
    line = list(color = "gray0", width = 2) # Set the color and width of the marker borders
  )
)
fig_3 <- fig_3 %>% layout(
  title = "Change in GDP vs. Interest Rate", # Set the plot title
  yaxis = list(zeroline = FALSE, title = "Interest Rate (% change)"), # Customize the Y-axis properties
  xaxis = list(zeroline = FALSE, title = "GDP (% change)") # Customize the X-axis properties
)
fig_3 # Display the scatter plot

# Scatter plot for GDP vs Tax
fig_4 <- plot_ly(
  data = final_data, # Specify the data source
  x = ~final_data$`GDP`, # Set the X-axis data from the "GDP" column
  y = ~final_data$`Tax`, # Set the Y-axis data from the "Tax" column
  marker = list(
    size = 5, # Set the size of the markers
    color = "red", # Set the color of the markers
    line = list(color = "gray0", width = 2) # Set the color and width of the marker borders
  )
)
fig_4 <- fig_4 %>% layout(
  title = "Change in GDP vs. Tax", # Set the plot title
  yaxis = list(zeroline = FALSE, title = "Tax (% change)"), # Customize the Y-axis properties
  xaxis = list(zeroline = FALSE, title = "GDP (% change)") # Customize the X-axis properties
)
fig_4 # Display the scatter plot


```


### Line Graphs 
```{r}
library(plotly)

# Line graph for Interest Rate by Country
interest_rate_graph <- plot_ly(final_data, x = ~countrycode, y = ~Int, type = 'scatter', mode = 'lines+markers', 
                               line = list(color = 'blue')) %>%
  layout(title = "Interest Rate by Country",
         xaxis = list(title = "Country"),
         yaxis = list(title = "Interest Rate"),
         showlegend = FALSE)

interest_rate_graph


```

```{r}
library(plotly)

# Line graph for Tax by Country
tax_graph <- plot_ly(final_data, x = ~countrycode, y = ~Tax, type = 'scatter', mode = 'lines+markers', 
                     line = list(color = 'red')) %>%
  layout(title = "Tax by Country",
         xaxis = list(title = "Country"),
         yaxis = list(title = "Tax"),
         showlegend = FALSE)

tax_graph

# Line graph for Unem (Unemployment) by Country
unem_graph <- plot_ly(final_data, x = ~countrycode, y = ~Unem, type = 'scatter', mode = 'lines+markers', 
                      line = list(color = 'green')) %>%
  layout(title = "Unemployment by Country",
         xaxis = list(title = "Country"),
         yaxis = list(title = "Unemployment Rate"),
         showlegend = FALSE)

unem_graph

# Line graph for GDP by Country
gdp_graph <- plot_ly(final_data, x = ~countrycode, y = ~GDP, type = 'scatter', mode = 'lines+markers', 
                     line = list(color = 'purple')) %>%
  layout(title = "GDP by Country",
         xaxis = list(title = "Country"),
         yaxis = list(title = "GDP"),
         showlegend = FALSE)

gdp_graph



```



```{r}
library(plotly)

# Trend line for Interest Rate (Int) by Year for different countries
fig <- plot_ly(final_data, x = ~year, y = ~Int, color = ~countrycode, type = "scatter", mode = "lines+markers",
               hoverinfo = 'text',
               text = ~paste("Year: ", year, "<br>Country: ", countrycode, "<br>Interest Rate: ", round(Int, 2))) %>%
  layout(showlegend = TRUE, xaxis = list(rangeslider = list(visible = TRUE)))

fig <- fig %>% config(
  scrollZoom = TRUE, editable = TRUE, selectdirection = "any", staticplot = FALSE
)

fig <- fig %>%
  layout(title = "<b>Trend in Interest Rates over Time by Country</b>",
         xaxis = list(title = "Year",
                      zerolinecolor = 'white',
                      zerolinewidth = 2.5,
                      gridcolor = 'white'),
         yaxis = list(title = "Interest Rate",
                      zerolinecolor = 'white',
                      zerolinewidth = 2.5,
                      gridcolor = 'white'),
         plot_bgcolor = 'lightblue', width = 950)

# Display the chart
fig
colnames(data)

```

### Line Charts - Trendlines 
```{r}
fig_tax <- plot_ly(final_data, x = ~year, y = ~Tax, type = "scatter", mode = "lines+markers",
                   color = ~countrycode, hoverinfo = 'text',
                   text = ~paste("Year: ", year, "<br>Country: ", countrycode, "<br>Tax: ", round(Tax, 2))) %>%
  layout(showlegend = TRUE, xaxis = list(rangeslider = list(visible = TRUE)))

fig_tax <- fig_tax %>% config(
  scrollZoom = TRUE, editable = TRUE, selectdirection = "any", staticplot = FALSE
)

fig_tax <- fig_tax %>%
  layout(title = "<b>Trend in Tax over Time by Country</b>",
         xaxis = list(title = "Year",
                      zerolinecolor = 'white',
                      zerolinewidth = 2.5,
                      gridcolor = 'white'),
         yaxis = list(title = "Tax",
                      zerolinecolor = 'white',
                      zerolinewidth = 2.5,
                      gridcolor = 'white'),
         plot_bgcolor = 'lightblue', width = 950)

fig_tax




```

```{r}
fig_gdp <- plot_ly(final_data, x = ~year, y = ~GDP, type = "scatter", mode = "lines+markers",
                   color = ~countrycode, hoverinfo = 'text',
                   text = ~paste("Year: ", year, "<br>Country: ", countrycode, "<br>GDP: ", round(GDP, 2))) %>%
  layout(showlegend = TRUE, xaxis = list(rangeslider = list(visible = TRUE)))

fig_gdp <- fig_gdp %>% config(
  scrollZoom = TRUE, editable = TRUE, selectdirection = "any", staticplot = FALSE
)

fig_gdp <- fig_gdp %>%
  layout(title = "<b>Trend in GDP per capital over Time by Country</b>",
         xaxis = list(title = "Year",
                      zerolinecolor = 'white',
                      zerolinewidth = 2.5,
                      gridcolor = 'white'),
         yaxis = list(title = "GDP",
                      zerolinecolor = 'white',
                      zerolinewidth = 2.5,
                      gridcolor = 'white'),
         plot_bgcolor = 'lightblue', width = 950)

fig_gdp

```

```{r}
fig_unem <- plot_ly(final_data, x = ~year, y = ~Unem, type = "scatter", mode = "lines+markers",
                    color = ~countrycode, hoverinfo = 'text',
                    text = ~paste("Year: ", year, "<br>Country: ", countrycode, "<br>Unemployment: ", round(Unem, 2))) %>%
  layout(showlegend = TRUE, xaxis = list(rangeslider = list(visible = TRUE)))

fig_unem <- fig_unem %>% config(
  scrollZoom = TRUE, editable = TRUE, selectdirection = "any", staticplot = FALSE
)

fig_unem <- fig_unem %>%
  layout(title = "<b>Trend in Unemployment over Time by Country</b>",
         xaxis = list(title = "Year",
                      zerolinecolor = 'white',
                      zerolinewidth = 2.5,
                      gridcolor = 'white'),
         yaxis = list(title = "Unemployment Rate",
                      zerolinecolor = 'white',
                      zerolinewidth = 2.5,
                      gridcolor = 'white'),
         plot_bgcolor = 'lightblue', width = 950)

fig_unem

```

## Pie Chart

```{r}
# Load necessary libraries
library(plotly)

# Create a blank graph
fig <- plot_ly()

# Add pie chart for Interest Rate (Int)
fig <- fig %>% add_pie(data = final_data, labels = ~countrycode, values = ~Int,
                       name = "Interest Rate", domain = list(row = 0, column = 0),
                       title = list(text = "Interest Rate"))

# Add pie chart for GDP
fig <- fig %>% add_pie(data = final_data, labels = ~countrycode, values = ~GDP,
                       name = "GDP", domain = list(row = 0, column = 1),
                       title = list(text = "GDP"))

# Add pie chart for Tax
fig <- fig %>% add_pie(data = final_data, labels = ~countrycode, values = ~Tax,
                       name = "Tax", domain = list(row = 1, column = 0),
                       title = list(text = "Tax"))

# Add pie chart for Unemployment (Unem)
fig <- fig %>% add_pie(data = final_data, labels = ~countrycode, values = ~Unem,
                       name = "Unemployment", domain = list(row = 1, column = 1),
                       title = list(text = "Unemployment"))

# Customize layout of the subplots
fig <- fig %>% layout(title = "Pie Charts with Subplots for Various Variables", showlegend = TRUE,
                      grid = list(rows = 2, columns = 2),
                      xaxis = list(showgrid = FALSE, zeroline = TRUE, showticklabels = TRUE),
                      yaxis = list(showgrid = FALSE, zeroline = TRUE, showticklabels = TRUE))

# Display the pie chart
fig

```

## Histogram

```{r}
# Load necessary libraries
library(plotly)

# Create a new plotly object with an alpha (transparency) value set to 0.6
fig <- plot_ly(alpha = 0.6)

# Add histograms for each variable
fig <- fig %>% add_histogram(x = final_data$Int, name = "Interest Rate")
fig <- fig %>% add_histogram(x = final_data$GDP, name = "GDP")
fig <- fig %>% add_histogram(x = final_data$Tax, name = "Tax")
fig <- fig %>% add_histogram(x = final_data$Unem, name = "Unemployment")

# Set the layout of the plotly object with "barmode" as "overlay" to overlay the histograms
fig <- fig %>% layout(barmode = "overlay",
                      title = "Overlayed Histograms for Various Variables",
                      xaxis = list(title = "Values"),
                      yaxis = list(title = "Frequency"))

# Display the overlaid histogram
fig

```

## Error Bars
```{r}
# Load necessary libraries
library(plyr)
library(plotly)

# Define new variables
x <- final_data$Unem   # Unemployment
y <- final_data$countrycode   # Country code
z <- final_data$Tax   # Tax

# Create a new dataset with x, y, and z variables
new_data <- data.frame(x, y, z)

# Calculate the mean output (Tax) for each combination of 'Unem' and 'countrycode'
data_4_mean <- ddply(new_data, c("y", "x"), summarise, Output = mean(z, na.rm = TRUE))

# Calculate the standard deviation of output (Tax) for each combination of 'Unem' and 'countrycode'
data_sd <- ddply(new_data, c("y", "x"), summarise, Output = sd(z, na.rm = TRUE))

# Combine the mean and standard deviation data into a new dataframe
dta_4 <- data.frame(data_4_mean, data_sd$Output)

# Rename the 'data_sd.Output' column to 'sd' for clarity
dta_4 <- rename(dta_4, c("data_sd.Output" = "sd"))

# Convert the 'x' column (Unemployment) to a factor for better plotting
dta_4$x <- as.factor(dta_4$x)

# Create a grouped bar plot using Plotly
fig <- plot_ly(data = dta_4[which(dta_4$y == "HND"),], x = ~x, y = ~Output, type = "bar", name = "HND",
               error_y = ~list(array = sd, color = "black"))
fig <- fig %>% add_trace(data = dta_4[which(dta_4$y == "DOM"),], name = "DOM")
fig <- fig %>% add_trace(data = dta_4[which(dta_4$y == "CRI"),], name = "CRI")
fig <- fig %>% add_trace(data = dta_4[which(dta_4$y == "PAN"),], name = "PAN")
fig <- fig %>% add_trace(data = dta_4[which(dta_4$y == "CAN"),], name = "CAN")
fig <- fig %>% add_trace(data = dta_4[which(dta_4$y == "MEX"),], name = "MEX")
fig <- fig %>% add_trace(data = dta_4[which(dta_4$y == "USA"),], name = "USA")

# Customize layout
fig <- fig %>% layout(title = "Mean Tax by Unemployment and Country with Standard Deviation",
                      xaxis = list(title = "Unemployment"),
                      yaxis = list(title = "Mean Tax"),
                      barmode = "group")

# Display the grouped bar plot
fig

```

## Violin Plots
```{r}
###violin plots

# Load necessary library
library(plotly)

# Create individual violin plots for each variable with year on the x-axis
fig_int <- plot_ly(final_data, x = ~year, y = ~Int, name = 'Interest Rate', type = 'violin',
                   box = list(visible = TRUE), meanline = list(visible = TRUE))

fig_gdp <- plot_ly(final_data, x = ~year, y = ~GDP, name = 'GDP', type = 'violin',
                   box = list(visible = TRUE), meanline = list(visible = TRUE))

fig_tax <- plot_ly(final_data, x = ~year, y = ~Tax, name = 'Tax', type = 'violin',
                   box = list(visible = TRUE), meanline = list(visible = TRUE))

fig_unem <- plot_ly(final_data, x = ~year, y = ~Unem, name = 'Unemployment', type = 'violin',
                    box = list(visible = TRUE), meanline = list(visible = TRUE))

# Combine the violin plots into a single subplot layout
fig <- subplot(fig_int, fig_gdp, fig_tax, fig_unem, nrows = 2, shareX = TRUE, shareY = TRUE)

# Customize the layout of the subplots
fig <- fig %>% layout(
  title = "Violin Plots for Various Metrics by Year",
  xaxis = list(
    title = "Year"
  ),
  yaxis = list(
    title = "Values",
    zeroline = FALSE
  )
)

# Display the subplot
fig


```

## contour Plot
```{r}
# Define x and y variables
x <- final_data$GDP  # GDP as the x variable
y <- final_data$Unem # Unemployment as the y variable

# Create a 2D density contour plot using plot_ly with x and y
fig <- plot_ly(x = x, y = y, type = "histogram2dcontour")

# Set layout options for the plot
layout_options <- list(
  showlegend = TRUE,
  margin = list(l = 80, r = 50, b = 50, t = 50),
  xaxis = list(title = "GDP"),
  yaxis = list(title = "Unemployment"),
  title = list(text = "2D Density Contour Plot of GDP vs. Unemployment")
)

# Update the layout options for the plot
fig <- fig %>% layout(layout_options)

# Display the plot
fig

```

```{r}
# Define x and y variables for Unemployment and Tax
x <- final_data$Unem  # Unemployment as the x variable
y <- final_data$Tax   # Tax as the y variable

# Create a 2D density contour plot using plot_ly with x and y
fig_unem_tax <- plot_ly(x = x, y = y, type = "histogram2dcontour")

# Set layout options for the plot
layout_options_unem_tax <- list(
  showlegend = FALSE,
  margin = list(l = 80, r = 50, b = 50, t = 50),
  xaxis = list(title = "Unemployment"),
  yaxis = list(title = "Tax"),
  title = list(text = "2D Density Contour Plot of Unemployment vs. Tax")
)

# Update the layout options for the plot
fig_unem_tax <- fig_unem_tax %>% layout(layout_options_unem_tax)

# Display the plot
fig_unem_tax

```

```{r}
# Define x and y variables for Unemployment and Interest Rate
x <- final_data$Unem  # Unemployment as the x variable
y <- final_data$Int   # Interest Rate as the y variable

# Create a 2D density contour plot using plot_ly with x and y
fig_unem_int <- plot_ly(x = x, y = y, type = "histogram2dcontour")

# Set layout options for the plot
layout_options_unem_int <- list(
  showlegend = FALSE,
  margin = list(l = 80, r = 50, b = 50, t = 50),
  xaxis = list(title = "Unemployment"),
  yaxis = list(title = "Interest Rate"),
  title = list(text = "2D Density Contour Plot of Unemployment vs. Interest Rate")
)

# Update the layout options for the plot
fig_unem_int <- fig_unem_int %>% layout(layout_options_unem_int)

# Display the plot
fig_unem_int

```

## Heatmap
```{r}
# Load necessary library
library(plotly)

# Subset the data to include only the relevant columns
data_subset <- final_data[, c("Int", "GDP", "Tax", "Unem", "Inf")]

# Calculate the correlation matrix
cor_matrix <- cor(data_subset, use = "complete.obs")  # use = "complete.obs" to handle NA values

# Create the heatmap
fig <- plot_ly(
  z = cor_matrix,
  type = "heatmap",
  colorscale = "Viridis",   # You can choose a different colorscale if desired
  x = colnames(cor_matrix),
  y = colnames(cor_matrix),
  text = round(cor_matrix, 2),
  hoverinfo = "text"
)

fig <- fig %>% config(
  scrollZoom = TRUE, editable = TRUE, selectdirection = "any", staticplot = FALSE
)

# Set axis labels and title
fig <- fig %>% layout(
  xaxis = list(title = "Variables"),
  yaxis = list(title = "Variables"),
  title = "Correlation Heatmap: Economic Indicators"
)

# Display the heatmap
fig

```

## Bubble Chart

```{r}
# Load necessary library
library(plotly)

# Filter the dataset to include only relevant columns for GDP, Unemployment, and Interest Rate
fuel1 <- final_data %>% select(GDP, Unem, Int)

# Create bubble chart
fig <- plot_ly(fuel1, x = ~GDP, y = ~Unem, type = "scatter", mode = "markers",
               color = I('blue'), size = ~Int,
               marker = list(sizes = c(5, 20), sizemode = "area", opacity = 0.5),
               text = ~paste("Interest Rate: ", Int),
               hoverinfo = "text") %>%
  layout(title = "GDP vs. Unemployment (Bubble Chart)",
         xaxis = list(title = "GDP"),
         yaxis = list(title = "Unemployment"))

# Add interactive features
fig <- fig %>% config(
  # Enable zooming functionality
  scrollZoom = TRUE,
  # Enable panning functionality
  editable = TRUE,
  # Enable data selection (filtering) by clicking and dragging on the plot
  selectdirection = "any",
  # Enable data selection (filtering) by holding down the shift key and clicking on the plot
  staticplot = FALSE
)

# Display the interactive plot
fig

```

